---
title: "README"
format: gfm
editor: visual
---

## ggbotbar

Create plot with geom_errorbar + geom_errorbarh with constant cap width.

## Usage

Load packages.

```{r load_pkg, message=FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 5,
  fig.dpi = 300
)

library(tidyverse)
library(ggbothbar)

packageVersion("ggbothbar")
```

```{r demo, message=FALSE}
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_errorbarb()
```

## Isotopic data workflows

The package includes helpers for mycoheterotrophic isotopic data. Start from replicate measurements, calculate enrichment, then visualise group-level uncertainty along both axes.

```{r iso-data, message=FALSE}
iso_raw <- tibble::tribble(
  ~type,       ~d13c, ~d15n,
  "reference", -20.1,   7.2,
  "reference", -19.8,   7.0,
  "sample",    -17.5,   9.8,
  "sample",    -18.0,   9.3
)

iso_enriched <- calc_enrichment(iso_raw, delta = c("d13c", "d15n"))
iso_enriched
```

### Error bars in both directions

`geom_errorbarb()` summarises each group into cross-shaped error bars. Supply a grouping aesthetic so replicates for the same taxon or treatment are combined.

```{r iso-barb, message=FALSE}
ggplot(
  iso_enriched,
  aes(d13c, d15n, colour = type, group = type, label = type)
) +
  geom_point(size = 1) +
  stat_mean_point(size = 2) +
  stat_mean_label(
    geom = "text",
    position = position_nudge(x = 0.25, y = 0.2),
    show.legend = FALSE
  ) +
  geom_errorbarb(fun.errorbar = "se", linewidth = 0.6, errorbar_tip_size = 1.2) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = label_isotope(13, "C"), y = label_isotope(15, "N"))
```

### Enrichment factor with Error boxes

```{r iso-box, message=FALSE}
ggplot() +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "grey50") +
  geom_vline(xintercept = 0, linetype = "dotted", colour = "grey50") +
  stat_mean_point(
    data = filter(iso_enriched, type != "reference"),
    mapping = aes(x = e13c, y = e15n),
    size = 2
  ) +
  geom_errorbarb(
    data = filter(iso_enriched, type != "reference"),
    mapping = aes(x = e13c, y = e15n),
    fun.errorbar = "sd"
  ) +
  geom_errorbox(
    data = filter(iso_enriched, type == "reference"),
    mapping = aes(x = e13c, y = e15n),
    fill = NA,
    fun.errorbar = "sd",
    colour = "darkgreen"
  ) +
  labs(x = label_isotope(13, "C", "epsilon"), y = label_isotope(15, "N", "epsilon")) +
  theme_aca(base_family = "Arial")
```
